\documentclass[unicode,11pt,a4paper,oneside,numbers=endperiod,openany]{scrartcl}

\renewcommand{\thesubsection}{\arabic{subsection}}
\usepackage{amsmath}


\input{assignment.sty}


\hyphenation{PageRank}
\hyphenation{PageRanks}


\begin{document}


\setassignment
\setduedate{Wednesday, 26 October 2022, 23:59 AM}

\serieheader{Numerical Computing}{2022}{Student: FULL NAME}{Discussed with: FULL NAME}{Solution for Project 2}{}
\newline

\assignmentpolicy


The purpose of this assignment\footnote{This document is originally 
based on a blog from Cleve Moler, who wrote a fantastic blog post about the Lake Arrowhead graph, and John 
Gilbert, who initially created the coauthor graph from the 1993 Householder Meeting. You can find more information 
at \url{http://blogs.mathworks.com/cleve/2013/06/10/lake-arrowhead-coauthor-graph/}.  Most of this assignment is derived
from this archived work.} is to learn the importance of sparse linear algebra algorithms to solve fundamental 
questions in social network analyses.
We will use the coauthor graph from the Householder Meeting and the social network of friendships from Zachary's karate club~\cite{karate}.
These two graphs are one of the first examples where matrix methods were used in computational social network analyses.

\section*{Social Networks [Total: 85 points + 15 points for report quality]}

\subsection{The Reverse Cuthill McKee Ordering [10 points]}

\subsection{Sparse Matrix Factorization [20 points]}

\subsubsection*{a)}

\[
    A_{ij} = 
    \begin{cases}
        1 & \text{if $i=1$ or $j=1$  or $j=n$ and $i \neq j$} \\
        n + i -1 & \text{if $i=j$}\\
        0 & \text{otherwise}
    \end{cases}
\]

% construct matrix:
% - A_{ij} = 1 if i=1 or j=1 or j=n and i != j
% - A_{ij} = n + i -1 if i=j
% - A_{ij} = 0 otherwise

\[A = 
\begin{bmatrix}
    1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 \\
    1 & 11 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
    1 & 0 & 12 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
    1 & 0 & 0 & 13 & 0 & 0 & 0 & 0 & 0 & 1 \\
    1 & 0 & 0 & 0 & 14 & 0 & 0 & 0 & 0 & 1 \\
    1 & 0 & 0 & 0 & 0 & 15 & 0 & 0 & 0 & 1 \\
    1 & 0 & 0 & 0 & 0 & 0 & 16 & 0 & 0 & 1 \\
    1 & 0 & 0 & 0 & 0 & 0 & 0 & 17 & 0 & 1 \\
    1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 18 & 1\\
    1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1
\end{bmatrix}
\]

Number of non-zero elements: 44;

\subsubsection*{b)}

A matrix $A \in R^{n \times n}$ will have $n \times n$ entries.
The non-zero entries are:
\begin{itemize}
    \item $n$ entries in the first row
    \item $n$ entries in the last row
    \item $n$ entries in the first column
    \item $n$ entries in the last column
    \item $n$ entries in the diagonal
\end{itemize}
So $5n$ minus the entries that are counted twice or more:
\begin{itemize}
    \item $A_{11}$ is counted in the first row, the first column and in the diagonal
    \item $A_{nn}$ is counted in the last row, the last column and in the diagonal
    \item $A_{1n}$ is counted in the first row and the last column
    \item $A_{n1}$ is counted in the last row and the first column
\end{itemize}
So the total number of non-zero entries is: $5n - 2 -2 -1-1 = 5n-6$.

\subsubsection*{d)}
% TODO: Julia code does not run

\subsubsection*{e)}



\subsection{Degree Centrality [5 points]}

\subsection{The Connectivity of the Coauthors [5 points]}

\subsection{PageRank of the Coauthor Graph [5 points]}

\subsection{Zachary's karate club: social network of friendships between 34 members [40 points]}

\begin{thebibliography}{99}
\bibitem{karate} The social network of a karate club at a US university, M.~E.~J. Newman and M. Girvan, Phys. Rev. E 69,026113 (2004)
pp. 219-229.
\end{thebibliography}


\end{document}